Using Model Checking to Analyze the System Behavior of the LHC Production Grid

### 2013-02-11 ElsevierFGCS_singleColumn.tex
library cleanup

To overcome the problem of state space explosion, several
variants to explicit model checking have gained popularity. Runtime
verification~\cite{citeulike:10198274,Havelund:2008:RVC:1424465.1424470}
aims to increase confidence in the correctness, but does not make claims for
absence of defects in a system.
The target program is instrumented manually, or automatically by scanning
function calls and variable assignments, 
in order to capture and emit relevant events which are then monitored against a
special component during execution.
This component resembles our monitoring process automaton used to cope with
state space explosion during model checking 
of DIRAC's WMS system. While runtime verification is useful for
partially-capturing incorrect behavior of the system implementation, 
code instrumentation is not always a feasible option, considering the increase
of memory consumption.

@@

Instead of constructing a model of the system from scratch, when the only 
source of documentation is the code itself, a technique called process (or workflow) mining~\cite{Agrawal:1998:MPM:645338.650397}
can be used. It is a method of automatically distilling a process or software specification from a set of 
structured traces collected from real system executions.
The mined specifications are typically expressed in a form of Message Sequence Graphs or Petri-Nets, 
graphic formalism variants of the classical finite-state machines. Given that the 
analysis of program execution traces is still the most prevalent debugging technique
for distributed systems, and DIRAC's components (as well as the middleware) produce
immense amounts of execution logs anyhow, it seems like an attractive approach to constructing models.
There is a vast amount of work in this area~\cite{Aalst:1997:VWN:647744.733919,Lou:2010:MPW:1835804.1835883,Walkinshaw:2008:IFM:1642931.1642962,vanderAalst:2004:WMD:1018031.1018329,Ammons:2002:MS:503272.503275,Kumar:2011:SMC:1985793.1986002}.
A common assumption in process mining is that each trace is totally ordered, 
which for concurrent systems does not always hold. To cope with this, most approaches 
require that the thread IDs or process IDs are recorded in the traces as well.
Observing this as a realistic obstacle for obtaining an accurate model, Lou et al. in~\cite{Lou:2010:MPW:1835804.1835883}
propose an algorithm for constructing models from traces of concurrent systems by expressing events as partially ordered.
Their algorithm identifies dependency relationships between pairs of events in interleaved
traces. Another common assumption is a direct result of the underlying techniques used for constructing the models, namely data mining 
and machine learning. These techniques consider ``erroneous activities'' or activities that
were not logged for some reason, as noise, and disregard trace sequences with no statistical significance.
As a result, a design flaw can be mistaken for a noise, or not captured at all if a particular 
execution path has a low probability of occurrence. A notable exception in this direction is~\cite{vanderAalst:2005:PMV:2099016.2099032}, where the emphasis is not only on process discovery,
but also on verification: desired or undesired properties can be expressed with Linear Temporal Logic, 
and the checker verifies whether the observed behavior in the traces matches the property.

>>>>

To overcome the problem of state space explosion, several
variants to explicit model checking have gained popularity. Runtime
verification~\cite{DBLP:conf/dagstuhl/ColinM04,Havelund:2008:RVC:1424465.1424470}
aims to increase confidence in the correctness, but does not make claims for
absence of defects in a system.
The target program is instrumented manually, or automatically by scanning
function calls and variable assignments, 
in order to capture and emit relevant events which are then monitored against a
special component during execution.
This component resembles our monitoring process automaton used to cope with
state space explosion during model checking 
of DIRAC's WMS system. While runtime verification is useful for
partially-capturing incorrect behavior of the system implementation, 
code instrumentation is not always a feasible option, considering the increase
of memory consumption.

@@

Instead of constructing a model of the system from scratch, when the only 
source of documentation is the code itself, a technique called process (or workflow) mining~\cite{Agrawal:1998:MPM:645338.650397}
can be used. It is a method of automatically distilling a process or software specification from a set of 
structured traces collected from real system executions.
The mined specifications are typically expressed in a form of Message Sequence Graphs or Petri-Nets, 
graphic formalism variants of the classical finite-state machines. Given that the 
analysis of program execution traces is still the most prevalent debugging technique
for distributed systems, and DIRAC's components (as well as the middleware) produce
immense amounts of execution logs anyhow, it seems like an attractive approach to constructing models.
There is a vast amount of work in this area~\cite{Aalst:1997:VWN:647744.733919,Lou:2010:MPW:1835804.1835883,Walkinshaw:2008:IFM:1642931.1642962,vanderAalst:2004:WMD:1018031.1018329,Ammons:2002:MS:503272.503275}.
A common assumption in process mining is that the events in each trace are totally ordered, 
which for concurrent systems does not always hold. To cope with this, most approaches 
require that the thread IDs or process IDs are recorded in the traces as well.
Observing this as a realistic obstacle for obtaining an accurate model, Lou et al. in~\cite{Lou:2010:MPW:1835804.1835883}
propose an algorithm for constructing models from traces of concurrent systems by expressing events as partially ordered.
Their algorithm identifies dependency relationships between pairs of events in interleaved
traces. Another common assumption is a direct result of the underlying techniques used for constructing the models, namely data mining 
and machine learning. These techniques consider ``erroneous activities'' or activities that
were not logged in the appropriate order for some reason, as noise, and disregard trace sequences with no statistical significance.
As a result, a design flaw can be mistaken for a noise, or not captured at all if a particular 
execution path has a low probability of occurrence. A notable exception in this direction is~\cite{vanderAalst:2005:PMV:2099016.2099032}, where the emphasis is not only on process discovery,
but also on verification: desired or undesired properties can be expressed with Linear Temporal Logic, 
and the checker verifies whether the observed behavior in the traces matches the property.
